opXRD: Open Experimental Powder X-ray Diffraction Database

### 2024-12-23 chapters/04_our_dataset.tex
Added general introduction to PCA

%TODO: A paragraph on how data is correlated
The degree of variation between the patterns is different for each contribution. For example, the CRNS and the INT contributions each are collections that encompass many research projects over a large period of time and hence exhibit a high degree of variability between individual patterns. In contrast, the contributions by the USC and the LBNL contain many very similar patterns. The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys collected on the same machine. The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes.

@@

@@

Another experimental dataset integrated with opXRD is a collection of powder diffraction data sourced from open-access publications and collaborating institutions. All collaborating institutions provided this data with full authorization for research purposes. This dataset provides broader coverage of chemical elements and includes ionic crystals, atomic crystals, and metallic crystals. This dataset surpasses Xred in size, containing 499 entries. However, contrary to Xred these data do not come paired with corresponding crystal structures.

>>>>

The below figure~$\eqref{fig:components}$ illustrates how varied the data of each contribution is using principle component analysis (PCA). PCA is a dimensional reduction technique that can be applied to datasets $X \subset \mathbb{R}^n$ to reduce the number of components needed to describe points $\p \in X$ up to some tolerance in lost accuracy. The cumulative explained variance ratio is a measure indicating how much of the variance in the dataset $X$ can be explained using a given number of components. For example, if PCA on a dataset $X \subset \mathbb{R}^10$ yields a cumulative variance ratio of 99\%, then each point in the dataset can be represented using only 5 degrees of freedom up to a minor loss in accuracy.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=\linewidth]{figures/component_fractions.png}
    \caption{Explained variance ratio over the fraction of the maximum number of components for each dataset contributed to opXRD. Each dataset is named after the contributing institutions. Datasets contributed by the same institution are enumerated alphabetically in the order in which they are described below. }
    \label{fig:components}
\end{figure}

%TODO: A paragraph on how data is correlated
The degree of variation between the patterns is different for each contribution. For example, the CRNS and the INT contributions each are collections that encompass many research projects over a large period of time and hence exhibit a high degree of variability between individual patterns. In contrast, the contributions by the USC and the LBNL contain many very similar patterns. The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys collected on the same machine. The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes. \\

@@

The following texts describe the datasets contributed by each collaborating institution. Each text includes a description of the investigated materials and how X-ray diffraction data was collected.

@@

Another experimental dataset integrated with opXRD is a collection of powder diffraction data sourced from open-access publications and collaborating institutions. All collaborating institutions provided this data with full authorization for research purposes. This dataset provides broader coverage of chemical elements and includes ionic crystals, atomic crystals, and metallic crystals. This dataset surpasses Xred in size, containing 499 entries. However, contrary to Xred these data do not come paired with corresponding crystal structures. \\
